{\pkg}: A Python Package for Integrating Computational Models Across Languages and Scales

### 2019-01-03 papers/lang2018/ms.tex
Added algorithms for async comms, connection drivers, model drivers, and master thread

\pdfoutput=1
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage[noend]{algpseudocode}
\ifdraft
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\else
		\usepackage{natbib}
		\citestyle{apj}
		\bibliographystyle{apj}
		\bibpunct[,]{(}{)}{;}{a}{}{,}
	\fi
\else
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\fi
\fi

@@

@@

%---------------------------------------------------------------------------------
% ASYNCHRONOUS MESSAGE MASSING
\subsubsection{Asynchronous Message Passing}\label{SSS:async}
%
The same asynchronous communication strategy is used with each of the communication mechanisms supported by {\pkg}. 
Models do not block on sending messages to 
output channels; the model is free to continue working on its task while 
an \emph{output driver} waits for the message to be routed and received on a separate master process thread. Similarly, 
an \emph{input driver} continuously pings input channels on another thread, moving received messages into a 
buffer queue so that they are ready and waiting for the receiving model 
when it asks for input. These drivers work together to move message along from one model to the next like a conveyor belt. 
As a result, models in complex integration networks are not affected by the rate at which dependent model consume their output and the speedup offered by running the models in parallel is improved. Figure \ref{fig:async} describes the general flow of messages.
%
\ifinclfig
 	\begin{figure}[htbp]
	\begin{center}
	\includegraphics[width=\columnwidth,keepaspectratio]{./images/io_drivers.png}
	\caption{Diagram of how messages are passed asynchronously using input/output drivers and an intermediate channel.}
	\label{fig:async}
	\end{center}
	\end{figure}
\fi
%
\begin{enumerate}
	\item A model sends a message in the form of a native data object via a language-specific API to one of the output channels declared in the model YAML. The output channel interface encodes the message and sends it.
	\item An output connection driver (written in Python) runs in a separate thread on the master process, listening to the model output channel. When the model sends a message, the output connection driver checks that the message is in the expected format and then forwards it to an intermediate channel. The intermediate channel is used as a buffer for support on distributed architectures (e.g. if one model is running on a remote machine). In these cases, the intermediate channel will connect to a RMQ broker using security credentials.
	\item An input connection driver (also written in Python) runs in a separate thread, listening to the intermediate channel. When a message is received, it is then forwarded to the input channel of the receiving model as specified in the integration network YAML.
	\item The receiving model receives the message in the form of an analogous native data object. Interface receive calls can be either blocking or non-blocking, but are blocking by default.
\end{enumerate}

@@

@@

\todo{algorithm?}

>>>>

\pdfoutput=1
\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
%\usepackage[noend]{algpseudocode}
\usepackage{caption}
\algblockdefx[Try]{Try}{EndTry}%
	{\textbf{try}}%
	{\textbf{end try}}
\algcblockdefx[TryExcept]{Try}{Except}{EndTry}%
	{\textbf{except}}%
	{\textbf{end try}}
\algblockdefx[class]{Class}{EndClass}%
	[2]{\textbf{class} {\sc #1}(#2)}%
	{\textbf{end class}}
\algblockdefx[classmethod]{ClassMethod}{EndClassMethod}%
	[2]{\textbf{method} {\sc #1}(#2)}%
	{\textbf{end method}}

%\DeclareCaptionFormat{algor}{%
%  \hrulefill\par\offinterlineskip\vskip1pt%
%    \textbf{#1#2}#3\offinterlineskip\hrulefill}
%\DeclareCaptionStyle{algori}{singlelinecheck=off,format=algor,labelsep=space}
%\captionsetup[algorithm]{style=algori}

\ifdraft
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\else
		\usepackage{natbib}
		\citestyle{apj}
		\bibliographystyle{apj}
		\bibpunct[,]{(}{)}{;}{a}{}{,}
	\fi
\else
	\ifieee
		\usepackage[numbers]{natbib}
		\bibliographystyle{ieeetr}
	\fi
\fi

@@

The algorithms for a generic model driver and the master thread can be found in Algorithms \ref{alg:model} \& \ref{alg:master} respectively.
% MODEL DRIVER
\begin{algorithm}[H]
    \caption{Generic model driver class. This algorithm does not represent the actual implementation of any one of the languages supported by {\pkg}. Each of the language-specific drivers will have slightly different methods for compilation (turning the source code into an executable), launching the model process, and monitoring the model's status via output and error codes. However, this algorithm does provided the general behavior pattern upon which every model driver is based.}
    \label{alg:model}
    \begin{algorithmic}[1]
    \Class{ModelDriver}{$name, args, yml$}
    	\State $errors \gets []$
    	\State $executable \gets$ \Call{Compile}{args}
	\State $process \gets$ \Call{ModelProcess}{} on a new process
	\State $thread \gets$ \Call{ModelThread}{} on a new thread
	\item[]
	%
	\ClassMethod{ModelProcess}{}
		\State \Call{Executable}{}
	\EndClassMethod
	\item[]
	%
	\ClassMethod{ModelThread}{}
		\While{\Call{IsAlive}{{\sc ModelProcess}}}
			\If{\Call{HasOutput}{{\sc ModelProcess}}}
				\State \textbf{Print} \Call{GetOutput}{{\sc ModelProcess}}
			\Else
				\State \textbf{Sleep}
			\EndIf
		\EndWhile
		\If{\Call{HasErrors}{{\sc ModelProcess}}}
			\State $errors \gets$ \Call{GetErrors}{{\sc ModelProcess}}
		\EndIf
	\EndClassMethod
	\item[]
	%
	\ClassMethod{Terminate}{}
		\State \Call{Kill}{{\sc ModelProcess}}
	\EndClassMethod
	\item[]
	%
	\ClassMethod{Stop}{}
		\While{not $errors$}
			\State \textbf{Sleep}
		\EndWhile
		\State \Call{Terminate}{}
	\EndClassMethod
	\item[]
    \EndClass
    \end{algorithmic}
\end{algorithm}

%
% MASTER THREAD
%\begin{center}
\begin{algorithm}[H]
	\caption{Basic steps executed on the master thread to create, start, and monitor model \& connection drivers. This pseudocode is not an exhaustive reproduction of the entire implementation, but represents the broad strokes needed to understand the purpose and behavior of the master thread. The algorithm for the {\sc ModelDriver} and {\sc ConnectionDriver} can be found in Algorithms \ref{alg:model} \& \ref{alg:conn} respectively.}
	\label{alg:master}
	\begin{algorithmic}[1]
	\State $models$ $\gets$ \textbf{Parse} YAML files
	\State $inputChannels \gets Map()$
	\State $outputChannels \gets Map()$
	% CREATE DRIVERS
	\item[] \Comment Create Drivers
	\ForAll{$m in models$}
		\ForAll{$i in m.inputs$}
			\If{not $i.isFile$}
				\State $inputAddress \gets i.address$
			\Else
				\State $inputAddress \gets `generate'$
			\EndIf
			\State $i.driver \gets$ \Call{ConnectionDriver}{$inputAddress, `generate', i$}
			\State $inputChannels[i.name] = i$
		\EndFor
	\EndFor
	\ForAll{$m in models$}
	\algstore{alg:master}
	\end{algorithmic}
\end{algorithm}
% PAGE BREAK
\begin{algorithm}[H]
	\caption{Continuation of Algorithm \ref{alg:master}}
	\begin{algorithmic}[1]
	\algrestore{alg:master}
		\ForAll{$o \in m.outputs$}
			\If{$o.isFile$}
				\State $outputAddress \gets o.address$
			\Else
				\If{$o.args \in inpurChannels$}
					\State $outputAddress \gets inputChannels[o.args]$
				\Else
					\State \Call{Error}{}
				\EndIf
			\EndIf
			\State $o.driver \gets$ \Call{ConnectionDriver}{$`generate', outputAddress, o$}
			\State $outputChannels[o.name] = o$
		\EndFor
	\EndFor
	\ForAll{$m in models$}
		\State $m.driver \gets$ \Call{ModelDriver}{$m.name, m.args, m$}
	\EndFor
	% START DRIVERS
	\item[] \Comment Start Drivers	
	\ForAll{$m in models$}
		\ForAll{$i in m.inputs$}
			\State $i.driver.$\Call{Start}{}
		\EndFor
	\EndFor
	\ForAll{$m in models$}
		\ForAll{$o \in m.outputs$}
			\State $o.driver.$\Call{Start}{}
		\EndFor
	\EndFor
	\ForAll{$m in models$}
		\State $m.driver.$\Call{Start}{}
	\EndFor
	% MONITOR DRIVERS
	\item[] \Comment Monitor Drivers
	\State $errorFlag \gets false$
	\While{not $errorFlag$}
		\ForAll{$m in models$}
			\If{$m.driver.errors$}
				\State $errorFlag \gets true$
				\State \textbf{break}
			\EndIf
		\EndFor
	\EndWhile
	% STOP DRIVER
	\item[] \Comment Stop Drivers
	\If{$errorFlag$}
		\State $allErrors \gets []$
		\ForAll{$m in models$}
			\State $m.$\Call{Terminate}{}
			\State \textbf{Append} $allErrors$, $m.errors$
		\EndFor
		\State \Call{Error}{$allErrors$}
	\Else
		\ForAll{$m in models$}
			\State $m.$\Call{Stop}{}
		\EndFor
	\EndIf
	\end{algorithmic}
\end{algorithm}
%\end{center}
%
\todo{make sure this is split correctly across pages}

@@

%---------------------------------------------------------------------------------
% ASYNCHRONOUS MESSAGE MASSING
\subsubsection{Asynchronous Message Passing}\label{SSS:async}
%
The same asynchronous communication strategy is used with each of the communication mechanisms supported by {\pkg}. 
Models do not block on sending messages to 
output channels; the model is free to continue working on its task while 
an \emph{output driver} waits for the message to be routed and received on a separate master process thread. Similarly, 
an \emph{input driver} continuously pings input channels on another thread, moving received messages into a 
buffer queue so that they are ready and waiting for the receiving model 
when it asks for input. These drivers work together to move message along from one model to the next like a conveyor belt. 
As a result, models in complex integration networks are not affected by the rate at which dependent model consume their output and the speedup offered by running the models in parallel is improved. Figure \ref{fig:async} describes the general flow of messages, Algorithms \ref{alg:async_send} \& \ref{alg:async_recv} describe the asynchronous output \& input channel procedures, and Algorithm \ref{alg:conn} describes the procedure followed by input \& output connection drivers.
%
\ifinclfig
 	\begin{figure}[htbp]
	\begin{center}
	\includegraphics[width=\columnwidth,keepaspectratio]{./images/io_drivers.png}
	\caption{Diagram of how messages are passed asynchronously using input/output drivers and an intermediate channel.}
	\label{fig:async}
	\end{center}
	\end{figure}
\fi
%
\begin{enumerate}
	\item A model sends a message in the form of a native data object via a language-specific API to one of the output channels declared in the model YAML. The output channel interface encodes the message and sends it.
	\item An output connection driver (written in Python) runs in a separate thread on the master process, listening to the model output channel. When the model sends a message, the output connection driver checks that the message is in the expected format and then forwards it to an intermediate channel. The intermediate channel is used as a buffer for support on distributed architectures (e.g. if one model is running on a remote machine). In these cases, the intermediate channel will connect to a RMQ broker using security credentials.
	\item An input connection driver (also written in Python) runs in a separate thread, listening to the intermediate channel. When a message is received, it is then forwarded to the input channel of the receiving model as specified in the integration network YAML.
	\item The receiving model receives the message in the form of an analogous native data object. Interface receive calls can be either blocking or non-blocking, but are blocking by default.
\end{enumerate}

% SENDER
\begin{algorithm}[H]
    \caption{Asynchronous communication class for sending messages to an output channel. The procedures {\sc Open, Close, Send, \& IsOpen} are procedures that are implemented by external communication packages and different slightly from one communication mechanism to the next.}
    \label{alg:async_send}
    \begin{algorithmic}[1]
    % SENDER
    \Class{AsyncSender}{$address$}
    	\State $backlog \gets []$
	\State $x \gets$ \Call{Open}{$address$}
	\State \textbf{Run} \Call{BacklogThread}{} on a new thread
	\item[]
	%
	\ClassMethod{BacklogThread}{}
		\While{\Call{IsOpen}{$x$}}
    			\If{$length(backlog) > 0$}
				\State $flag \gets$ \Call{Send}{$x$, $backlog[0]$}
				\If{$flag$}
					\State \textbf{Remove} $backlog[0]$
				\Else
					\State \Call{Close}{$x$}
				\EndIf
			\Else
				\State \textbf{Sleep}
			\EndIf
    		\EndWhile
    	\EndClassMethod
	\item[]
	%
	\ClassMethod{SendBacklog}{$msg$}
		\If{\Call{IsOpen}{$x$}}
			\State \textbf{Append} $backlog$, $msg$
		\EndIf
		\State \textbf{Return} \Call{IsOpen}{$x$}
	\EndClassMethod
	\item[]
	%
	\ClassMethod{Close}{}
		\State \Call{Close}{$x$}
		\State $backlog \gets []$
	\EndClassMethod
	\item[]
    \EndClass
    \end{algorithmic}
\end{algorithm}

% RECEIVER
\begin{algorithm}[H]
    \caption{Asynchronous communication class for receiving messages from an input channel. The procedures {\sc Open, Close, Recv, IsOpen, \& IsMessageWaiting} are procedures that are implemented by external communication packages and different slightly from one communication mechanism to the next.}
    \label{alg:async_recv}
    \begin{algorithmic}[1]
    \item[]
    \Class{AsyncReceiver}{$address$}
    	\State $backlog \gets []$
	\State $x \gets$ \Call{Open}{$address$}
	\State $thread \gets$ \Call{BacklogThread}{} on a new thread
	\item[]
	%
	\ClassMethod{BacklogThread}{}
		\While{\Call{IsOpen}{$x$}}
    			\If{\Call{IsMessageWaiting}{$x$}}
				\State $flag, msg \gets$ \Call{Recv}{$x$}
				\If{$flag$}
					\State \textbf{Append} $backlog$, $msg$
				\Else
					\State \Call{Close}{$x$}
				\EndIf
			\Else
				\State \textbf{Sleep}
			\EndIf
		\EndWhile
	\EndClassMethod
	\item[]
	%
	\ClassMethod{RecvBacklog}{$msg$}
		\If{$length(backlog) > 0$}
			\State $msg \gets backlog[0]$
			\State \textbf{Remove} $backlog[0]$
			\State \textbf{Return} true, $msg$
		\Else
			\State \textbf{Return} \Call{IsOpen}{$x$}, ''
		\EndIf
	\EndClassMethod
	\item[]
	%
	\ClassMethod{Close}{}
		\State \Call{Close}{$x$}
		\State $backlog \gets []$
	\EndClassMethod
	\item[]
    \EndClass
    \end{algorithmic}
\end{algorithm}
%

%
\begin{algorithm}[H]
    \caption{Connection Driver}
    \label{alg:conn}
    \begin{algorithmic}[1]
    \Class{ConnectionDriver}{$inputAddress, outputAddress, yml$}
    	\State $input \gets$ \Call{AsyncReceiver}{$inputAddress$}
    	\State $output \gets$ \Call{AsyncSender}{$outputAddress$}
	\State $thread \gets$ \Call{ConnectionThread}{} on a new thread
	\item[]
	%
	\ClassMethod{ConnectionThread}{}
		\While{true}
			\State $flag, msg \gets$ $input.$\Call{RecvBacklog}{}
			\If{not $flag$} \State \textbf{break} \EndIf
			\If{$msg$}
				\State \textbf{Transform} $msg$ based on units and expected data format
				\State $flag \gets$ $output.$\Call{SendBacklog}{$msg$}
				\If{not $flag$} \State \textbf{break} \EndIf
			\Else
				\State \textbf{Sleep}
			\EndIf
		\EndWhile
		%
		\State $input.$\Call{CloseAsync}{}
		\State $output.$\Call{CloseAsync}{}
	\EndClassMethod
	\item[]
	%
	\ClassMethod{Terminate}{}
		\State $input.$\Call{CloseAsync}{}
		\State $output.$\Call{CloseAsync}{}
	\EndClassMethod
	\item[]
	%
	\ClassMethod{Stop}{}
		\While{\Call{IsOpen}{$input.x$} and \Call{IsOpen}{$output.x$}}
			\State \textbf{Sleep}
		\EndWhile
		\State \Call{Terminate}{}
	\EndClassMethod
	\item[]
    \EndClass
    \end{algorithmic}
\end{algorithm}

\todo{discuss algorithm}

@@

%\begin{algorithm}[H]
%    \caption{Large Messages}
%    \label{alg:large}
%    \begin{algorithmic}[1] 
%    \end{algorithmic}
%\end{algorithm}

@@
